---
title: "Supervised learning in high dimension: R"
author: "Sophie Dabo"
output:
  beamer_presentation: 
    keep_tex: yes
  ioslides_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, include = FALSE}
ozone <- read.table("/Users/sophiedabo/Nextcloud/Cours_big_data_inference/ozone.txt",header=T,sep="")
```

## Linear Regression: Pollution Example

Before beginning the analysis, it is helpful to start by visualizing the data.

```{r}
library(knitr)
kable(head(ozone), format = "markdown", digits = 2)
```

Correlation

```{r}
library(corrplot)
library(GGally)
X=ozone[,2:12]
matcor=cor(X)
corrplot(matcor, type="upper", tl.col="black", tl.srt=45)
#ggpairs(X, columns = 1:8, aes(alpha = 0.8))
```

```{r, echo = TRUE}
library("scatterplot3d")
scatterplot3d(ozone[,"T12"],ozone[,"Vx12"],ozone[,"maxO3"],type="h", pch=16,box=FALSE,xlab="Tempéature à 12h",ylab="Vent",zlab="Ozone")

```

## Data

```{r, include = TRUE}
head(ozone)
ozoneframe=as.data.frame(ozone[,-1])
```

PCA

```{r}
library("FactoMineR")
library("factoextra")
res.pca=PCA(ozoneframe,quali.sup = c(12,13))
fviz_pca_biplot(res.pca, select.ind = list(contrib = 50), 
                select.var = list(contrib = 50),ggtheme = theme_minimal(),habillage=12)
```


## Example of pollution

We consider a linear model between ozone, wind, and cloudiness.

```{r, echo = TRUE}
modele1 <- lm(maxO3~Vx12+T12,data=ozone)
summary(modele1)
anova(modele1)
```

## Model without intercept

```{r, include = TRUE}
model1<- lm(maxO3~-1+Vx12+T12,data=ozone)
pred=predict(model1)
summary(model1)
```

## Without wind

```{r, include = TRUE}
model1<- lm(maxO3~Ne12,data=ozone)
summary(model1)
model2<- lm(maxO3~Vx9+Vx12+Vx15+Ne9+Ne12+Ne15+T9+T12+T15+maxO3v,data=ozone)
summary(model2)
```

## Exhaustive search

The "leaps" package allows you to perform an exhaustive search using the "regsubsets" function. For more than 50 variables, the option "really.big=TRUE" must be added to avoid excessive computation time.

```{r, include = TRUE}
library(leaps)
choix <- regsubsets(maxO3~Vx9+Vx12+Vx15+Ne9+Ne12+Ne15+T9+T12+T15+maxO3v,int=T,nbest=1,nvmax=11,method="exh", data=ozone)
resume=summary(choix)
resume
ozone03 <- as.numeric(rownames(resume$wh))+1
```

##  R2 Choice

Keep  Vx9, Ne9, T12, maxO3v

```{r,echo = TRUE}
plot(choix,scale="adjr2")
```



## By BIC

Keep Intercept Vx9, Ne9, T12, maxO3v

```{r,echo = TRUE}
plot(choix,scale="bic")
```

## If  p>>n?

Let us consider the data on the contents of fat, sugar, flour, and water in biscuits (Osbone et al., 1984) for 72 biscuits, available in the "ppls" package. For these 72 biscuits, the compositions in fat, sugar, flour, and water are measured using a classical approach, while the spectrum is observed over all wavelengths between 1100 and 2498 nanometers, regularly spaced every 2 nanometers. We therefore have 700 observed values, or potentially explanatory variables, for each biscuit dough sample.

Typically, this study is conducted in a very high-dimensional context with $p >> n$. On these uncooked biscuits, absorbance is measured for wavelengths between 1100 and 2498 nanometers, regularly spaced every 2 nanometers. There are thus 700 potentially explanatory variables.

Let us consider the percentage of sugars. We have $p = 700$ variables for $n = 40$ individuals. The classical least squares estimator is not defined.

 
The classical least squares estimator is not defined.

## Lectrure des données


```{r, echo = TRUE}
library(ppls)
data(cookie)
# extraire le taux de sucrose et les spectres
cook = data.frame(cookie$constituents$sucrose,cookie$NIR)
names(cook)= c("sucrose",paste("X",1:700,sep=""))
```


## Data

```{r,echo = TRUE}
# Response
hist(cook[,"sucrose"])
```


```{r,echo = TRUE}
# spectres by sucrose intensity
coul=rainbow(20)[as.integer(as.factor
    (as.integer(cook[,1]-10)))]
ts.plot(t(cook[,-1]),col=coul) 
```

## MCO estimation not valide


```{r,echo = TRUE}
cook.app=cook[1:40,]
summary(lm(sucrose ~ ., data=cook.app))
```

## Extract the training and test samples

```{r,echo = TRUE}
cook.app=cook[1:40,]
cook.test=cook[41:72,]
```

## Ridge regression

Ici on a les paramètres $\beta$ en fonction de la  régularisation

```{r, include=TRUE, echo=TRUE}
library(MASS)
 plot(lm.ridge(sucrose  ~ ., data=cook.app,
   lambda = seq(0,1,0.01)))
```

## Régression ridge avec plusieurs paramètres de régularisation


```{r, include=TRUE, echo=TRUE}
 plot(lm.ridge(sucrose  ~ ., data=cook.app,
   lambda = seq(0,0.2,0.001)))
cook.ridge=lm.ridge(sucrose  ~ ., data=cook.app,
   lambda = seq(0,0.2,0.001))
```

## Choix du paramètre de régularisation par validation croisée

The $n=40$ observations in the training sample are randomly divided into 4 subsamples of size 10.

```{r,echo = TRUE, include=TRUE}
library(pls)
set.seed(87)
cvseg=cvsegments(nrow(cook.app),k=4,type="random")
```

## Selection of the regularization parameter by cross-validation

For $\tau$ (here $\lambda$) between $0$ and $1$, we estimate $\beta$ and the corresponding PRESS on each subsample. We sum the 4 PRESS values to obtain the cross-validation PRESS and choose the $\tau$ that minimizes this PRESS.

```{r,echo = TRUE}
choix.kappa=function(kappamax,cvseg,nbe=100){
press=rep(0,nbe)
for (i in 1:length(cvseg)){
valid=cook.app[unlist(cvseg[i]),]
modele=lm.ridge(sucrose~.,
   data=cook.app[unlist(cvseg[-i]),],
lambda=seq(0,kappamax,length=nbe))
coeff=coef(modele)
prediction=matrix(coeff[,1],nrow(coeff),
   nrow(valid))+coeff[,-1]%*%
   t(data.matrix(valid[,-1]))
press=press+rowSums((matrix(valid[,1],
nrow(coeff),nrow(valid),byrow=T)-
   prediction)^2)
      }
kappaet=seq(0,kappamax,length=
   nbe)[which.min(press)]
return(list(kappaet=kappaet,press=press))
}

```

## Creation of a function to plot residuals versus predicted Y values

```{r,echo = TRUE}
plot.res=function(x,y,titre="title")
{
plot(x,y,col="blue",ylab="Residuals",
xlab="Predicted values",main=titre)
abline(h=0,col="green")
}

```

## Display of results with the optimal PRESS

```{r,echo = TRUE}
# execution

res=choix.kappa(0.5,cvseg,nbe=1000)
plot(seq(0,0.5,length=1000),res$press,ylab="PRESS",
xlab="lambda") # optimal value

```

## Ridge regression with the optimal parameter

```{r,echo = TRUE}
kappaet=res$kappaet
cook.ridgeo=lm.ridge(sucrose~.,data=cook.app,
lambda=kappaet)
coeff=coef(cook.ridgeo)

# Computation of predicted responses

fit.rid=rep(coeff[1],nrow(cook.app))+
as.vector(coeff[-1]%*%t(data.matrix(cook.app[,-1])))

```

## Plot of predictions versus true values

```{r,echo = TRUE}
# True responses versus predicted responses

plot(fit.rid,cook.app[,"sucrose"],ylab="True responses",
xlab="Predicted responses")

# The line y = x

lines(cook.app[,"sucrose"],cook.app[,"sucrose"])

```

## Residuals

```{r,echo = TRUE}
# Computation of residuals

res.rid=fit.rid-cook.app[,"sucrose"]

# Residuals versus predicted responses

plot.res(fit.rid,res.rid,titre="")

```

## Prediction on the test sample and computation of prediction error

Since the data are centered and scaled, the appropriate transformation must be applied to obtain the estimated response.

```{r,echo = TRUE}
# prediction on the test sample

ychap=rep(coeff[1],nrow(cook.test))+
as.vector(coeff[-1]%*%t(data.matrix(cook.test[,-1])))

```

## Prediction on the test sample and computation of prediction error

```{r,echo = TRUE}
plot(ychap,cook.test[,1],ylab="True responses",
xlab="Predicted responses")

## Prediction error

mean((cook.test[,1]-ychap)^2)

```

## Comparison between Ridge and OLS coefficients


```{r,echo = TRUE}
# Linear model by OLS: lambda = 0

coefflm <- coef(lm.ridge(sucrose~.,data = cook.app,
lambda=0))

```

## Continuation

```{r,echo = TRUE}
# Plot of coefficients from both models

matplot(t(rbind(coeff,coefflm)),type="l",col=c(2,1))
legend("topleft", col=c(2,1), lty=c(1,2),
legend=c("OLS beta", "Ridge beta"), cex=0.8)

```

## LASSO: Selection of the regularization parameter

We choose 1000 values of $\lambda$ and search, via cross-validation (on the mean squared error = MSE, using lars), for the optimal value. The plot shows the mean squared error as a function of the constraint fraction relative to $\lambda$, $\frac{\lambda}{|\hat{\beta}|_1}$.

```{r,echo = TRUE}
# A sequence of lambda values

library(lars)
lambdas <- seq(from = 0, to = 1, length = 1000)

```

## Continuation

```{r,echo = TRUE}
# Optimal lambda or delta by CV using lars

mse.cv <- cv.lars(data.matrix(cook.app[,-1]), cook.app[,1],K = 4,se=F,index=lambdas,use.Gram=F)
lambdas.op <-lambdas[which.min(mse.cv$cv)]

```

## Prediction using the optimal parameter


```{r,echo = TRUE}
cook.lasso=lars(data.matrix(cook.app[,-1]),
   cook.app[,1],use.Gram=F)
fit.lasso=predict(cook.lasso,
   data.matrix(cook.app[,-1]),s=lambdas.op,
mode="fraction")$fit
```

## Prediction using the optimal parameter


```{r,echo = TRUE}
# Plot predictions

plot(fit.lasso,cook.app[,"sucrose"], ylab="True responses",
xlab="Predicted responses")
res.lasso=fit.lasso-cook.app[,"sucrose"]

```

## Residuals using the optimal parameter


```{r,echo = TRUE}
# Plot predictions versus residuals

plot.res(fit.lasso,res.lasso)

```

## Prediction error using the optimal parameter


```{r,echo = TRUE}
# Mean squared prediction error

pred.lasso=predict(cook.lasso,
data.matrix(cook.test[,-1]),
s=lambdas.op,mode="fraction")$fit
mean((pred.lasso-cook.test[,"sucrose"])^2)

```

## Regression PCR

We choose the number of components by cross-validation.

```{r,echo = TRUE}
# Mean squared errors

library(pls)
cook.pcr=pcr(sucrose~.,ncomp=28,data=cook.app,
validation="CV",segments=cvseg)
msepcv.pcr=MSEP(cook.pcr,estimate= c("train","CV"))

```

## PCR regression

Mean errors as a function of the number of components.

```{r,echo = TRUE}
plot(msepcv.pcr,type="l")
ncompo=which.min(msepcv.pcr$val["CV",,])-1
ncompo
```

## Regression PCR

Evolution of the proportion of variance of the variables explained by each component.

```{r,echo = TRUE}
# Mean squared errors

cook.pcro=pcr(sucrose~.,ncomp=ncompo,
data=cook.app)

```

## Regression PCR

```{r,echo = TRUE}
# Mean squared errors

plot(explvar(cook.pcro),type="l",main="", pin=0.5)

```

## Regression PCR

We compute the prediction error on the test sample.

```{r,echo = TRUE}
fit.pcr=predict(cook.pcro,ncomp=1:6)[,,6]
res.pcr=fit.pcr-cook.app[,"sucrose"]

```

## Prediction error on the test sample

```{r,echo = TRUE}
plot.res(fit.pcr,res.pcr)
ychap=predict(cook.pcro,newdata=
cook.test)[,1,ncompo]
mean((cook.test[,"sucrose"]-ychap)^2)

```

## Regression PLS

PLS regression with up to 28 orthogonal components and residual plot with 28 components.
We choose the number of components by cross-validation.

```{r,echo = TRUE}
cook.pls=plsr(sucrose~.,ncomp=28,
data=cook.app,validation="CV",segments=cvseg)
print(cook.pls)
```

## Regression PLS

```{r,echo = TRUE}
plot(cook.pls)
msepcv.pls=MSEP(cook.pls,estimate=
   c("train","CV"))
```

## Regression PLS

We plot the evolution of the mean squared error as a function of the number of PLS components.

```{r,echo = TRUE}
plot(msepcv.pls,type="l")
```

## Regression PLS

We plot the evolution of the proportion of variance of the variables explained by each component.


```{r,echo = TRUE}
plot(explvar(cook.pls),type="l",main="")
```

## PLS optimal by cross-validation

```{r}
ncompo=which.min(msepcv.pls$val["CV",,])-1
ncompo

# optimal model

cook.plso=plsr(sucrose~.,ncomp=ncompo,
data=cook.app)

# residuals

fit.pls=predict(cook.plso,ncomp=1:ncompo)[,,ncompo]
plot(fit.pls,cook.app[,"sucrose"],fin=0.5)
res.pls=fit.pls-cook.app[,"sucrose"]
plot.res(fit.pls,res.pls)

```

##Prediction error on the test sample

PLS is better than PCR.

```{r,echo = TRUE}
# mean squared error

ychap=predict(cook.plso,newdata=
cook.test)[,1,ncompo]
plot(ychap,cook.test[,1])
mean((cook.test[,"sucrose"]-ychap)^2)

```


